### 2024-02-12 outputs/paper/paper.qmd
update folder

---
title: "Grade Distribution Among Genders"
subtitle: "My subtitle if needed"
author: 
  - Faiza Imam 
  - Quang Mai 
  - Catherine Punnoose 
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Load packages
library(readstata13)
library(readr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(janitor)

# Load data
# inst_data <- read.csv(here::here("inputs/data/instsurvey.csv"))
std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))
# exp_data <- read.csv(here::here("inputs/data/experiment.csv"))
```

# TESTING GRAPHS

Our data (@tbl-studentgender).

```{r}
#| label: tbl-studentgender
#| tbl-cap: "Number of students enrolled in class, by gender"
#| warning: false
#| message: false
#| echo: false

std_data |>
  count(std_female) %>% 
  kable(
    col.names = c("Gender", "Number of students")
  )
```

# Introduction

<intro>

# Data {#sec-data}

<data>

# Model
<model>

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References