### 2024-02-12 testing_graphs.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
---

```{r}
#| include: false
#| warning: false
#| message: false

#loading packages
library(readstata13)
library(readr)

#reading in the stata file
instsurvey <- read.dta13(here::here("inputs/data/instsurvey.dta"))
stdsurvey <- read.dta13(here::here("inputs/data/stdsurvey.dta"))
experiment <- read.dta13(here::here("inputs/data/experiment.dta"))

#writing to csv
write_csv(instsurvey, here::here("inputs/data/instsurvey.csv"))
write_csv(stdsurvey, here::here("inputs/data/stdsurvey.csv"))
write_csv(experiment, here::here("inputs/data/experiment.csv"))

```

# TESTING GRAPHS
```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)

# inst_data <- read.csv(here::here("inputs/data/instsurvey.csv"))
std_data <- read.csv(here::here("inputs/data/stdsurvey.csv"))
# exp_data <- read.csv(here::here("inputs/data/experiment.csv"))

std_data |>
  count(std_standing)

```

```{r}
#| include: false
#| warning: false
#| message: false

# Install and load necessary packages
install.packages("knitr")
install.packages("kableExtra")
library(knitr)
library(kableExtra)

# Create a data frame with the given values
data <- data.frame(
  Category = c("Percent of records", "Grade change", "Positive grade change", 
               "Negative grade change", "Female instructor", "Days between grade changes", 
               "GPA", "Percent in College of Agriculture", "College of Business", 
               "College of Engineering", "College of Human Sciences", 
               "College of Liberal Arts", "College of Natural Resources", 
               "College of Natural Sciences", "College of Veterinary Sciences", 
               "Intra-university", "Observations"),
  All = c(100, 0.464, 0.439, 0.025, 44.9, 25.8, 3.04, 4.5, 10.8, 4.8, 
          12.6, 32.2, 4.6, 23.3, 3.4, 3.9, 1341552),
  Female = c(53.4, 0.424, 0.404, 0.021, 50.0, 24.9, 3.13, 4.9, 8.6, 1.9, 
              15.5, 33.6, 3.5, 23.2, 4.4, 4.4, 716772),
  Male = c(46.6, 0.509, 0.479, 0.031, 38.9, 26.7, 2.94, 4.0, 13.2, 8.2, 
            9.1, 30.6, 5.9, 23.4, 2.3, 3.3, 624780)
)

# Create a kable table
tbl <- kable(data, format = "html", align = "c") %>%
  kable_styling()

# Print the table
print(tbl)

```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows.

# Data {#sec-data}

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References