### 2024-02-12 outputs/paper/paper.qmd
clean up folder

---
title: "Grade Distribution Among Genders"
subtitle: "My subtitle if needed"
author: 
  - Faiza Imam 
  - Quang Mai 
  - Catherine Punnoose 
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Load packages
library(readstata13)
library(readr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(janitor)
library(dplyr)

# Load data
inst_data <- read.csv(here::here("outputs/data/inst_data_cleaned.csv"))

std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))

# exp_data <- read.csv(here::here("inputs/data/experiment.csv"))
```

# FIGURE 3

Our data (@tbl-figure3).

```{r}
#| label: tbl-figure3
#| tbl-cap: "Instructor Survey: Regrade Results by Timing and Students’ Gender"
#| warning: false
#| message: false
#| echo: false

#Code Assisted from Chat GPT 3.5
# End of Semester Results MALE

male_dta <- inst_data |>
              select(maleup_fn, malesame_fn, maledown_fn)

maleup_avg <- mean(male_dta$maleup_fn, na.rm = TRUE)
malesame_avg <- mean(male_dta$malesame_fn, na.rm = TRUE)
maledown_avg <- mean(male_dta$maledown_fn, na.rm = TRUE)

males <- data.frame(
  value = c(maleup_avg, malesame_avg, maledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

# Define the desired order of categories
category_order <- c("Grade Decrease", "No Change", "Grade Increase")

# Convert the 'category' variable to a factor with the specified order
males$category <- factor(males$category, levels = category_order)

# Set custom breaks for the y-axis
custom_breaks <- seq(0, 90, by = 30)

# Create a bar graph using ggplot
ggplot(males, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(title = "Panel A. End of semester, male students",
       x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

# End of Semester Results FEMALE

female_dta <- inst_data |>
              select(femaleup_fn, femalesame_fn, femaledown_fn)

femaleup_avg <- mean(female_dta$femaleup_fn, na.rm = TRUE)
femalesame_avg <- mean(female_dta$femalesame_fn, na.rm = TRUE)
femaledown_avg <- mean(female_dta$femaledown_fn, na.rm = TRUE)

females <- data.frame(
  value = c(femaleup_avg, femalesame_avg, femaledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

# Define the desired order of categories
category_order <- c("Grade Decrease", "No Change", "Grade Increase")

# Convert the 'category' variable to a factor with the specified order
females$category <- factor(females$category, levels = category_order)

# Set custom breaks for the y-axis
custom_breaks <- seq(0, 90, by = 30)

# Create a bar graph using ggplot
ggplot(females, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(title = "Panel A. End of semester, female students",
       x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal()

```

# FIGURE 2

Figure 2. Student Survey: Number of Classes Students Considered for Regrade Requests by Students’ Gender

```{r}
classes_dta <- std_data |>
              select(std_male, num_class, consider_regrade)|>
              rename(Gender = std_male) |>
              filter(is.na(consider_regrade) | consider_regrade != "No")

classes_dta <- classes_dta |>
  group_by(num_class, Gender, consider_regrade) |>
  count() |>
  group_by(Gender, consider_regrade) |>
  mutate(Percentage = round(n/sum(n)*100, 1))|>
  mutate(num_class = case_match(num_class, 
                             0 ~ "0 class",
                             1 ~ "1 class",
                             2 ~ "2 classes",
                             3 ~ "3 classes",
                             4 ~ "4 classes",
                             5 ~ "5 classes"))

classes_dta$consider_regrade[classes_dta$consider_regrade == "Yes"] <- "No"
classes_dta$consider_regrade[is.na(classes_dta$consider_regrade)] <- "Yes"
classes_dta$Gender[classes_dta$Gender == "1"] <- "Male"
classes_dta$Gender[classes_dta$Gender == "0"] <- "Female"

classes_dta |>
  filter(consider_regrade != "No") |>
  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade)

head(classes_dta, 10) |>
  kable(
    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
    booktabs = TRUE,
    align = c("c", "c", "c", "c", "c")
  )

classes_dta |>
ggplot(aes(x = num_class, y = Percentage, fill = Gender)) + 
  geom_bar(stat="identity", position = "dodge") +
  theme_minimal() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  geom_text(aes(label=Percentage), position=position_dodge(width=0.9), vjust=-0.25) +
  labs(x = "Classes",
       y = "Distribution by gender (percent)") +
  scale_fill_manual(values = c("Male" = "#4B659D", "Female" = "#A83B48")) +
  scale_x_discrete(labels = c("0 class", "1 class", "2 classes", "3 classes", "4 classes", "5+ classes"))

```

# Introduction

<intro>

# Data {#sec-data}

<data>

# Model
<model>

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References